---
title: "Appendix"
author: "Harrison Magee, Michael Feron, Blake Waechter"
date: "December 9, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.
Below, we gather and process our data
```{r prelim, echo = TRUE, results='hide'}
bodyfat.dat = read.table("http://inside.mines.edu/~wnavidi/math424/project/bodyfat.dat", 
                         header = TRUE)
bodyfatDF = as.data.frame(bodyfat.dat)

min(bodyfatDF$Bodyfat)
bodyfatDF <- bodyfatDF[-c(which.min(bodyfatDF$Bodyfat)),]
min(bodyfatDF$Bodyfat)
max(bodyfatDF$Bodyfat)

min(bodyfatDF$Age)
max(bodyfatDF$Age)

min(bodyfatDF$Weight)
max(bodyfatDF$Weight)

min(bodyfatDF$Height)
bodyfatDF <- bodyfatDF[-c(which.min(bodyfatDF$Height)),]
min(bodyfatDF$Height)
max(bodyfatDF$Height)

min(bodyfatDF$BMI)
max(bodyfatDF$BMI)

min(bodyfatDF$Abdomen)
max(bodyfatDF$Abdomen)

min(bodyfatDF$Ankle)
max(bodyfatDF$Ankle)

min(bodyfatDF$Biceps)
max(bodyfatDF$Biceps)

min(bodyfatDF$Chest)
max(bodyfatDF$Chest)

min(bodyfatDF$Forearm)
max(bodyfatDF$Forearm)

min(bodyfatDF$Hip)
max(bodyfatDF$Hip)

min(bodyfatDF$Knee)
max(bodyfatDF$Knee)

min(bodyfatDF$Neck)
max(bodyfatDF$Neck)

min(bodyfatDF$Thigh)
max(bodyfatDF$Thigh)

min(bodyfatDF$Wrist)
max(bodyfatDF$Wrist)

Bodyfat = bodyfatDF$Bodyfat
Age = bodyfatDF$Age
Weight = bodyfatDF$Weight
Height = bodyfatDF$Height
BMI = bodyfatDF$BMI
Abdomen = bodyfatDF$Abdomen
Ankle = bodyfatDF$Ankle
Biceps = bodyfatDF$Biceps
Chest = bodyfatDF$Chest
Forearm = bodyfatDF$Forearm
Hip = bodyfatDF$Hip
Knee = bodyfatDF$Knee
Neck = bodyfatDF$Neck
Thigh = bodyfatDF$Thigh
Wrist = bodyfatDF$Wrist
```

## 2.
Boxplots for the data to help verify data cleaning process
```{r box, echo = FALSE, out.width='.49\\textwidth', fig.height = 5, fig.width=4}
boxplot(Bodyfat, main="Bodyfat")
boxplot(Age, main="Age")
boxplot(Weight, main="Weight")
boxplot(Height, main="Height")
boxplot(BMI, main="BMI")
boxplot(Abdomen, main="Abdomen")
boxplot(Ankle, main="Ankle")
boxplot(Biceps, main="Biceps")
boxplot(Chest, main="Chest")
boxplot(Forearm, main="Forearm")
boxplot(Hip, main="Hip")
boxplot(Knee, main="Knee")
boxplot(Neck, main="Neck")
boxplot(Thigh, main="Thigh")
boxplot(Wrist, main="Wrist")
```

## 3.
Below we fit a simple linear model
```{r lm_all, echo = TRUE, results = 'show'}
all.lm = lm(Bodyfat ~ Age + Weight + Height + BMI + Neck + Chest + Abdomen + Hip + 
              Thigh + Knee + Ankle + Biceps + Forearm + Wrist)
all.sum = summary(all.lm)
all.sum
```

## 4.
Scater plots for high p-val variables
```{r p-testing, echo=FALSE, out.width='\\textwidth', fig.height = 4}
plot(Height, Bodyfat, main = "Bodyfat Vs. Height")
plot(BMI, Bodyfat, main = "Bodyfat Vs. BMI")
plot(Chest, Bodyfat, main = "Bodyfat Vs. Chest")
plot(Knee, Bodyfat, main = "Bodyfat Vs. Knee")
plot(Ankle, Bodyfat, main = "Bodyfat Vs. Ankle")
plot(Biceps, Bodyfat, main = "Bodyfat Vs. Biceps")
```

## 5.
Fit a linear model without height
```{r model_without_height, echo = TRUE, results='show'}
new1.lm = lm(Bodyfat ~ Age + Weight + BMI + Neck + Chest + Abdomen + Hip + Thigh
             + Knee + Ankle + Biceps + Forearm + Wrist)
new1.sum = summary(new1.lm)
new1.sum
```

## 6.
New linear model dropping Ankle, Chest, Knee, Biceps
```{r model_without_Ankle_Chest_Knee_Biceps, echo = TRUE, results='show'}
new2.lm = lm(Bodyfat ~ Age + Weight + BMI + Neck + Abdomen + Hip + Thigh 
             + Forearm + Wrist)
new2.sum = summary(new2.lm)
new2.sum
```

## 7.
Calculate $F$-statistic for dropping Ankle, Chest, Knee, Biceps
```{r f_new3, echo=TRUE, results='show'}
anova(new2.lm, new1.lm)
```

## 8.
Investigating the colinearity of BMI
```{r BMI_colinearity, echo = FALSE, fig.align = "center", fig.height=4}
plot(BMI, Weight, main = "Weight vs BMI")
plot(BMI, Height, main = "Height vs BMI")
plot(BMI, 1/(Height*Height), main="1/Height^2 vs BMI")
```

## 9.
New model dropping BMI
```{r model_without_BMI, echo = TRUE}
new3.lm = lm(Bodyfat ~ Age + Weight + Neck + Abdomen + Hip + Thigh + Forearm + 
               Wrist)
new3.sum = summary(new3.lm)
new3.sum
```

## 10.
Running best subsets on full linear model
```{r best_subsets_full, echo = TRUE}
library("leaps")
bestall = regsubsets(Bodyfat ~ Age + Weight + Height + BMI + Neck + Chest + 
                       Abdomen + Hip + Thigh + Knee + Ankle + Biceps + Forearm + 
                       Wrist, data = bodyfatDF[,2:15], nbest = 1, nvmax=15)
bestall.sum = summary(bestall)
bestall.sum$which
cbind(bestall.sum$rsq, bestall.sum$adjr2, bestall.sum$cp)
```

## 11.
Calculate $F$ statistic for the new model having dropped all insignificant variables
```{r best_model, echo=TRUE}
anova(new3.lm, all.lm)
```

## 12.
Plotting the variables against Bodyfat to help identify tranformations to use
```{r transform_plots, echo=TRUE, out.width='\\textwidth', fig.height = 4}
plot(Age, Bodyfat, main='Bodyfat vs. Age')
plot(Weight, Bodyfat, main='Bodyfat vs. Weight')
plot(Neck, Bodyfat, main='Bodyfat vs. Neck')
plot(Abdomen, Bodyfat, main='Bodyfat vs. Abdomen')
plot(Hip, Bodyfat, main='Bodyfat vs. Hip')
plot(Thigh, Bodyfat, main='Bodyfat vs. Thigh')
plot(Forearm, Bodyfat, main='Bodyfat vs. Forearm')
plot(Wrist, Bodyfat, main='Bodyfat vs. Wrist')
```

## 13.
Calculating the linear models for all interactions of Body Measurements
```{r interactions, echo=TRUE}
NeckAbdomen = Neck*Abdomen
NeckHip = Neck*Hip
NeckThigh = Neck*Thigh
NeckForearm = Neck*Forearm
NeckWrist = Neck*Wrist
AbdomenHip = Abdomen*Hip
AbdomenThigh = Abdomen*Thigh
AbdomenForearm = Abdomen*Forearm
AbdomenWrist = Abdomen*Wrist
HipThigh = Hip*Thigh
HipForearm = Hip*Forearm
HipWrist = Hip*Wrist
ThighForearm = Thigh*Forearm
ThighWrist = Thigh*Wrist
ForearmWrist = Forearm*Wrist
Neck2 = Neck^2
Abdomen2 = Abdomen^2
Hip2 = Hip^2
Thigh2 = Thigh^2
Forearm2 = Forearm^2
Wrist2 = Wrist^2
logNeck = log(Neck)
logAbdomen = log(Abdomen)
logHip = log(Hip)
logThigh = log(Thigh)
logForearm = log(Forearm)
logWrist = log(Wrist)

interactions.lm = lm(Bodyfat ~ Age + Weight + Neck + Abdomen + Hip + Thigh + 
                       Forearm + Wrist + NeckAbdomen + NeckHip + NeckThigh + 
                       NeckForearm + NeckWrist + AbdomenHip + AbdomenThigh + 
                       AbdomenForearm + AbdomenWrist + HipThigh + HipForearm + 
                       HipWrist + ThighForearm + ThighWrist + ForearmWrist + 
                       Neck2 + Abdomen2 + Hip2 + Thigh2 + Forearm2 + Wrist2 + 
                       logNeck + logAbdomen + logHip + logThigh + logForearm + 
                       logWrist)

interactions.sum = summary(interactions.lm)
interactions.sum
```

## 14.
Best Subsets on the interactions
```{r best_subsets_interactions, echo = TRUE}
interactDF = as.data.frame(cbind(Age, Weight, Neck, Abdomen, Hip, Thigh, Forearm, 
                                 Wrist, NeckAbdomen, NeckHip, NeckThigh, NeckForearm, 
                                 NeckWrist, AbdomenHip, AbdomenThigh, AbdomenForearm, 
                                 AbdomenWrist, HipThigh, HipForearm, HipWrist, 
                                 ThighForearm, ThighWrist, ForearmWrist, Neck2, 
                                 Abdomen2, Hip2, Thigh2, Forearm2, Wrist2, logNeck, 
                                 logAbdomen, logHip, logThigh, logForearm, logWrist))


interactbest = regsubsets(Bodyfat ~ Age + Weight + Neck + Abdomen + Hip + Thigh + 
                            Forearm + Wrist + NeckAbdomen + NeckHip + NeckThigh + 
                            NeckForearm + NeckWrist + AbdomenHip + AbdomenThigh + 
                            AbdomenForearm + AbdomenWrist + HipThigh + HipForearm + 
                            HipWrist + ThighForearm + ThighWrist + ForearmWrist + 
                            Neck2 + Abdomen2 + Hip2 + Thigh2 + Forearm2 + Wrist2 + 
                            logNeck + logAbdomen + logHip + logThigh + logForearm + 
                            logWrist, data = interactDF, nbest = 1, nvmax=35)
interactbest.sum = summary(interactbest)
interactbest.sum$which
cbind(bestall.sum$rsq, bestall.sum$adjr2, bestall.sum$cp)
maxradj = which.max(bestall.sum$adjr2)
mincp = which.min(bestall.sum$cp)
maxradj
mincp
```

## 15.
Fit linear model for the best subsets
```{r best_subsets, echo=TRUE}
subsets.lm = lm(Bodyfat ~ Age + Weight + Thigh + Neck + Abdomen + Forearm + 
                  Wrist + NeckAbdomen + NeckThigh + ForearmWrist + Abdomen2 + 
                  Forearm2)
subsets.sum = summary(subsets.lm)
subsets.sum
```

## 16.
Stepwise Regression on all interactions
```{r stepwise_interactions, echo=TRUE}
model0 = lm(Bodyfat ~ 1, data = interactDF)
model1 = lm(Bodyfat ~ ., data = interactDF)
step(model0, scope = list(lower=model0, upper=model1), direction="both")
```

## 17.
We fit the linear model suggested by stepwise regression
```{r stepwise, echo=TRUE}
stepwise.lm = lm(Bodyfat ~ logAbdomen + HipWrist + Neck + ThighForearm + 
                   Age + Abdomen + Hip + Wrist + Thigh + Forearm)
stepwise.sum = summary(stepwise.lm)
stepwise.sum
```

## 18.
Compare the linear models
```{r compare, echo = TRUE, results='show'}
new3.sum
subsets.sum
stepwise.sum
```

## 19.
Test our model with training and testing data
```{r train_test, echo=TRUE}
newDF_x = as.data.frame(cbind(Age, Weight, Neck, Abdomen, Hip, Thigh, 
                              Forearm, Wrist))
train = sample(1:250, 200)
test = (-train)
trainX = newDF_x[train,]
trainY = Bodyfat[train]
testX = newDF_x[test,]
testY = Bodyfat[test]

train.lm = lm(trainY ~ Age + Weight + Neck + Abdomen + Hip + Thigh + Forearm + 
                Wrist, data=trainX)
train.sum = summary(train.lm)
train.sum

test.lm = lm(testY ~ Age + Weight + Neck + Abdomen + Hip + Thigh + Forearm + 
               Wrist, data=testX)
test.sum = summary(test.lm)
test.sum
```

## 20.
Plotting residuals for the model and residuals for the single variables
```{r residuals, echo=TRUE, fig.height=4}
plot(new3.lm)
Residuals = new3.lm$residuals
plot(Age, Residuals, main="Residuals vs. Age")
plot(Weight, Residuals, main="Residuals vs. Weight")
plot(Neck, Residuals, main="Residuals vs. Neck")
plot(Abdomen, Residuals, main="Residuals vs. Abdomen")
plot(Hip, Residuals, main="Residuals vs. Hip")
plot(Thigh, Residuals, main="Residuals vs. Thigh")
plot(Forearm, Residuals, main="Residuals vs. Forearm")
plot(Wrist, Residuals, main="Residuals vs. Wrist")
```

## 21.
Removing the high leverage points from the model and refitting the model
```{r lev, echo=TRUE, results='show'}
# new data model includes age, weight, neck, abdomen, hip, thigh, forearm, wrist
# without the two high leverage points
bodyfatDF_no_lev <- bodyfatDF[-c(39, 174),]
Bodyfat_no_lev = bodyfatDF_no_lev$Bodyfat
Age_no_lev = bodyfatDF_no_lev$Age
Weight_no_lev = bodyfatDF_no_lev$Weight
Neck_no_lev = bodyfatDF_no_lev$Neck
Abdomen_no_lev = bodyfatDF_no_lev$Abdomen
Hip_no_lev = bodyfatDF_no_lev$Hip
Thigh_no_lev = bodyfatDF_no_lev$Thigh
Forearm_no_lev = bodyfatDF_no_lev$Forearm
Wrist_no_lev = bodyfatDF_no_lev$Wrist

newDF_x_no_lev = as.data.frame(cbind(Age_no_lev, Weight_no_lev, Neck_no_lev, 
                                     Abdomen_no_lev, Hip_no_lev, Thigh_no_lev, 
                                     Forearm_no_lev, Wrist_no_lev))
train_no_lev = sample(1:248, 200)
test_no_lev = (-train_no_lev)
trainX_no_lev = newDF_x_no_lev[train_no_lev,]
trainY_no_lev = Bodyfat_no_lev[train_no_lev]
testX_no_lev = newDF_x_no_lev[test_no_lev,]
testY_no_lev = Bodyfat_no_lev[test_no_lev]

train.lm_no_lev = lm(trainY_no_lev ~ Age_no_lev + Weight_no_lev + Neck_no_lev
                     + Abdomen_no_lev + Hip_no_lev + Thigh_no_lev + Forearm_no_lev 
                     + Wrist_no_lev, data=trainX_no_lev)
train.sum_no_lev = summary(train.lm_no_lev)
train.sum_no_lev

test.lm_no_lev = lm(testY_no_lev ~ Age_no_lev + Weight_no_lev + Neck_no_lev
                    + Abdomen_no_lev + Hip_no_lev + Thigh_no_lev + Forearm_no_lev + 
                      Wrist_no_lev, data=testX_no_lev)
test.sum_no_lev = summary(test.lm_no_lev)
test.sum_no_lev
```



